Growth and inequality in public good provision --- an extended replication

### 2022-11-03 analysis/quarto/paper.tex
many things happened here...

\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\journal{Journal of Behavioral and Experimental Economics}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-harv}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Growth and inequality in public good provision \textbar{} an extended replication},
  pdfauthor={Hauke Roggenkamp},
  pdfkeywords={Replication study, Non-convenience sample, Open
science, Dynamic public good game, Online experiment, Generalizability},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{frontmatter}
\title{Growth and inequality in public good provision \textbar{} an
extended replication}
\author[1,2]{Hauke Roggenkamp%
%
}
 \ead{hauke.roggenkamp@unisg.ch}

@@

\begin{abstract}
This is the abstract. Lorem ipsum dolor sit amet, consectetur adipiscing
elit. Vestibulum augue turpis, dictum non malesuada a, volutpat eget
velit. Nam placerat turpis purus, eu tristique ex tincidunt et. Mauris
sed augue eget turpis ultrices tincidunt. Sed et mi in leo porta
egestas. Aliquam non laoreet velit. Nunc quis ex vitae eros aliquet
auctor nec ac libero. Duis laoreet sapien eu mi luctus, in bibendum leo
molestie. Sed hendrerit diam diam, ac dapibus nisl volutpat vitae.
Aliquam bibendum varius libero, eu efficitur justo rutrum at. Sed at
tempus elit.
\end{abstract}

\begin{keyword}
    Replication study \sep Non-convenience sample \sep Open
science \sep Dynamic public good game \sep Online experiment \sep 
    Generalizability
\end{keyword}
\end{frontmatter}\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, frame hidden, breakable, sharp corners, enhanced, interior hidden]}{\end{tcolorbox}}\fi

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

\begin{quote}
``There are two possible articles you can write: (a) the article you
planned to write when you designed your study or (b) the article that
makes the most sense now that you have seen the results. They are rarely
the same, and the correct answer is (b).'' \citep[p.~171]{bemwriting}
\end{quote}

\hypertarget{methodology}{%
\section{Methodology}\label{methodology}}

In the terminology of \citet{Hamermesh2007}, I ran both a \emph{pure} as
well as a \emph{scientific replication} of one treatment arm of
\citet{GMTV2017}'s dynamic public good game. The pure replication
re-analyzes the original data.
\protect\hyperlink{A:-Pure-Replication}{Appendix A} documents
\href{}{errors I identified} in the original paper. The scientific
replication, where I utilize a different sample drawn from a different
population in a different situation, is described in this section.

\hypertarget{procedure}{%
\subsection{Procedure}\label{procedure}}

Participants entered the experiment at appointed times remotely from
home. They first saw a welcome screen. After agreeing to the privacy
policy, they could proceed to the instructions individually. Having read
these instructions, each participant has also seen a demo-screen
explaining the user interface. Before proceeding, they had to answer six
comprehension questions correctly. Subsequently, they saw a waiting
screen until they could be matched with three other participants, who
have answered the comprehension questions correctly. Once matched, they
were exposed to the decision screen over ten periods. At the end of the
last period, participants saw results of all periods. Subsequently, they
were exposed to a voluntary climate action, where they could donate
(some of) their earnings to offset carbon dioxide. Subsequently, I
elicited risk preferences \citep{HoltLaury2002} and finished with
\citet{GMTV2017}'s questionnaire.

\hypertarget{experimental-design}{%
\subsection{Experimental Design}\label{experimental-design}}

As in the NOPUNISH 10 Period treatment arm of \citet{GMTV2017}, I ran
sessions with groups of four (\(i \in I=\{1,2,3,4\}\)), an initial
endowment of \(N_i^1 = 20\) tokens, \(T=10\) rounds, a private account
with a return of \(1\) and a group account with a return of \(1.5\)
(\(\Rightarrow\) MPCR\(\equiv \frac{1.5}{4}\)), such that:

\[
N_i^{t+1}=N_i^t - c_i^t + \frac{1.5}{4}\sum_{j=1}^4 c_j^t
\] Instead of receiving fresh endowments every period, participants
receive one initial endowment only at the beginning of the first period.
A participant's subsequent endowment equals her profit from the current
period such that a decision in one period has consequences on her future
endowments. For this reason, the game is described as a \emph{dynamic}
public good game

\hypertarget{sample}{%
\subsection{Sample}\label{sample}}

I recruited the participants from the so called
\emph{\href{https://www.wiso.uni-hamburg.de/forschung/forschungslabor/umfragelabor/aktuelle-umfragen/hamburgpanel.html}{HamburgPanel}}
using HROOT \citep{hroot}. The panel is provided by the University of
Hamburg's Research Laboratory, which used a randomized last digits
approach to build the panel while drawing from the population of
citizens from Hamburg, Germany.

@@

\hypertarget{software}{%
\subsection{Software}\label{software}}

The experiment was created using oTree \citep{oTree} and can be found on
\href{https://github.com/Howquez/coopUncertainty}{GitHub}.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{online-feasibility}{%
\subsection{Online Feasibility}\label{online-feasibility}}

How did the participants, who have never participated in a online group
experiment before, cope with the situation?

@@

{\centering \includegraphics{paper_files/figure-pdf/plotTime-1.pdf}

@@

\caption{Average Time Spent for each Contribution per Period}

@@

\hypertarget{pre-registered-gmtv-replication}{%
\subsection{Pre-registered GMTV
Replication}\label{pre-registered-gmtv-replication}}

\textbf{Result 1.} \emph{The \texttt{NOPUNISH\ 10} treatment of
\citet{GMTV2017} can be replicated because the replication data resemble
the original data with respect to initial and final contributions,
wealth and growth as well as inequality between and within groups.}

This is remarkable given the different sample and language, the
different software and user interface as well as the online setting
during the COVID19 pandemic.

\hypertarget{contributions}{%
\subsubsection{Contributions}\label{contributions}}

First, I ask whether the samples differ with respect to their initial
contributions to the public good. Is our replication sample more
pro-social than the original sample? Figure 1 reveals that it is not.
The distributions of both samples look fairly similar. Both samples
contributed 10 tokens, that is, 50\% of their endowments on average
(median and mean).\footnote{The two-sided rank sum test (comparing
  differences between samples) yields a p-Value of 0.3926 for the mean
  contribution in first round of the game.} Moreover, both samples'
initial contributions resemble initial contributions participants
usually make in the standard game with partner matching.\footnote{See
  Figure 3B in \citet{fehrgaechter2000} (p.989), for instance.} However,
in the dynamic game presented here, we are particularly interested in
the subsequent periods because differences add up exponentially. Do the
two groups remain similar over the course of time?

@@

{\centering \includegraphics{paper_files/figure-pdf/firstRoundViz-1.pdf}

@@

\caption{Individual contributions to the dynamic public good in the
first period}

In particular, do the two samples' contributions follow the same path
over the 10 periods they played? The answer is \emph{no}. Figure 2
illustrates that the samples make similar contributions at the beginning
and the end of the game but behave differently in between. More
precisely, the left panel--depicting the average contributions in
absolute terms--shows that the original sample contributed substantially
more than the replication sample \emph{in all but the first and last
periods}. For this reason, the original sample's behavior differs from
the replication sample's behavior in two aspects: it contributes more
and exhibits a considerable drop in the last period (whereas the
replication sample's contributions flatten).

Note that increasing contributions over time imply increasing endowments
over time. Hence, absolute contributions do not us much about the
willingness to cooperate. For this reason, the right panel in Figure 2
shows the average \emph{share of endowments contributed} over time. Both
samples exhibit a similar pattern: they decline and do not stabilize.
However, both samples also differ with respect to one aspect: the
replication sample's share of contributions declines faster.

{\centering \includegraphics{paper_files/figure-pdf/plotShareOfContributions-1.pdf}

\caption{The average amount of tokens contributed over time in
treatments.}

@@

\hypertarget{wealth-creation}{%
\subsubsection{Wealth Creation}\label{wealth-creation}}

How do the different contribution-paths translate into
wealth?\footnote{To measure wealth and growth, I define a variable
  called \emph{stock} which sums the endowments of all participants in a
  given group at the end of the round (that is, after the contributions
  have been made, multiplied and redistributed).} Given that the
original sample contributed more in most of the periods, one would
expect the respective groups to be considerably more wealthy. Figure 3
indicates just that. The grey lines show that an average group in the
original sample accumulated about 478 tokens. In contrast, an average
group in the replication sample accumulated about 380 tokens. This
difference is insignificant at conventional levels\footnote{The
  two-sided rank sum test (comparing differences between samples) yields
  a p-Value of 0.1356 for the mean stock in last round of the game.}
though.

@@

{\centering \includegraphics{paper_files/figure-pdf/stockDistributionViz-1.pdf}

@@

\caption{Groups' income at the end of the game}

@@

Although there clearly is growth, groups do not realize the maximal
potential efficiency: under full cooperation, a group can accumulate at
least 4613 tokens or EUR 230. This is depicted in the left panel of
Figure 4, where one can see the average wealth over time by sample. The
panel illustrates for both samples that growth was continuous and
surprisingly linear, given the exponential character of the game's
design. To sum up, the contribution behavior differed between samples.
In contrast, neither the eventual wealth nor the corresponding growth
paths differed. Differences in contribution behavior did, thus, not
translate in significantly different wealth outcomes.

Why? Perhaps because the heterogeneity within samples and across groups
has been too large to \emph{detect} a significant difference. The right
panel of Figure 4 depicts heterogeneity: In the replication sample, the
richest group earned 1425 tokens (which is about 1781\% of the initial
endowment) whereas the poorest group ends up with 92 tokens (115\%).
More broadly, the replication sample is characterized by inequality
between groups (\(SD_{Replication} =\) 336.06). The same holds true for
the original sample (\(SD_{Original} =\) 393.58). Hence, the
heterogeneity across groups does not differ between samples, which is
remarkable because the replication sample was drawn from a more
heterogeneous (non-convenience sample). Does it differ within groups?

@@

{\centering \includegraphics{paper_files/figure-pdf/growthHeterogeneityViz-1.pdf}

@@

\caption{Average wealth over time across samples.}

@@

\hypertarget{inequality}{%
\subsubsection{Inequality}\label{inequality}}

Given the different samples and the possibility of endogenous
growth--which essentially is the main feature of the game--I ask whether
and how the inequality grows \emph{within} groups. Figure 5 illustrates
that inequality did grow: at the end of the game, the original and the
replication groups exhibit an average Gini coefficient of 0.23 and 0.22,
respectively.\footnote{The two-sided rank sum test (comparing
  differences between samples) yields a p-Value of 0.6059 for the mean
  Gini coefficient in last round of the game.} Because every participant
started with the same initial endowment (in \emph{Period 0}, so to
speak), every group started equally--with a Gini coefficient equaling
zero.

Figure 6 shows that this initial state of equality ended with the first
period already: both samples exhibit a stark incline in inequality
before the second period started. From then on, the respective Gini
coefficients grew slowly but continuously -- for both samples.

@@

{\centering \includegraphics{paper_files/figure-pdf/giniDistributionViz-1.pdf}

@@

\caption{Groups' Gini coefficients (within groups) at the end of the
game}

@@

{\centering \includegraphics{paper_files/figure-pdf/unnamed-chunk-1-1.pdf}

@@

\caption{Average Gini coefficient (within groups) over time across
samples}

@@

\hypertarget{generalizability}{%
\subsection{Generalizability}\label{generalizability}}

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

\hypertarget{appendix}{%
\section{Appendix}\label{appendix}}

\hypertarget{a-pure-replication}{%
\subsection{A: Pure Replication}\label{a-pure-replication}}

\hypertarget{b-growth-and-inequality-exploratory}{%
\subsection{\texorpdfstring{B: Growth \emph{and} Inequality
(exploratory)}{B: Growth and Inequality (exploratory)}}\label{b-growth-and-inequality-exploratory}}

In contrast to \citet{GMTV2017}, I did not ask how rich and poor groups
differ. Instead, I was wondering, whether equal groups are wealthier.
More precisely, does the Gini coefficient correlate with growth and
wealth creation? To answer that question, Figure 7 applies a median
split showing equal and unequal groups' wealth over time.

>>>>

\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\journal{Journal of Behavioral and Experimental Economics}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-harv}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Growth and inequality in public good provision --- an extended replication},
  pdfauthor={Hauke Roggenkamp},
  pdfkeywords={Replication study, Non-convenience sample, Open
science, Dynamic public good game, Online experiment, Generalizability},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{frontmatter}
\title{Growth and inequality in public good provision --- an extended
replication}
\author[1,2]{Hauke Roggenkamp%
%
}
 \ead{hauke.roggenkamp@unisg.ch}

@@

\begin{abstract}
You can find the most recent version of this paper
\href{https://github.com/Howquez/coopUncertainty/blob/main/analysis/quarto/paper.pdf}{here}.
The abstract follows at later point.
\end{abstract}

\begin{keyword}
    Replication study \sep Non-convenience sample \sep Open
science \sep Dynamic public good game \sep Online experiment \sep 
    Generalizability
\end{keyword}
\end{frontmatter}\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, sharp corners, frame hidden, boxrule=0pt, breakable, enhanced, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi

\hypertarget{sec-intro}{%
\section{Introduction}\label{sec-intro}}

Today's actions are tomorrow's result. There are many settings in which
current decisions affect future outcomes and with it, future decision
spaces. Opting for environmental friendly policies today not only
reduces carbon dioxide omissions but also helps us to reach the Paris
climate targets tomorrow. Deferring these policies, the targets can be
reached as well---but with less flexibility. Hence, today's actions (or
the omission thereof) not only affect intermediate results but also the
number of paths one can choose to reach certain goals.

Standard public good games---although often intended to inform climate
policies\footnote{See \citet[p.~1]{GKLS2020} for numerous references.}---miss
these temporal interdependencies because participants have the same set
of actions in each period. A game implemented by Gächter, Mengel, Tsakas
\& Vostroknutov \citeyearpar{GMTV2017} (hereafter, GMTV) as well as
Stefan Große (2011, unpublished) incorporated interdependencies into a
\emph{dynamic} public good game with students in Nottingham, England and
Erfurt, Germany.

Because their setting is more realistic, one wonders whether it is
better suited to inform public policy. To answer this question, I
replicated parts of GMTV's experiment with different samples. In
addition, I observed the participant's behavior in voluntary climate
actions (VCA). This yielded a setting similar to \citet{GKLS2020}'s
which allows me to analyze how behavior in the abstract game translates
into real-world action across samples.

I utilized an inexperienced sample that has not been exposed to
interactive experiments before. Furthermore, I conducted the experiment
online. This required me to make the experiment's software robust to
minimize attrition. Because I observed metrics to assess the fluency and
feasibility of online experiments, this study also serves as a case
study like the one of \citet{AGM2018}.

Taken together, this study makes three contributions. First, it
replicates parts of GMTV's original experiment and highlights the
importance of pure replications. Second, it shows that logistically
complex online experiments are feasible for samples other than students
or clickworkers. Third, it shows that findings from abstract games do
not generalize well and even worse with the representative sample. After
reporting the methods, this paper is organized along these findings.

\hypertarget{sec-methods}{%
\section{Methodology}\label{sec-methods}}

In the terminology of \citet{Hamermesh2007}, I ran both a \emph{pure} as
well as a \emph{scientific replication} of one treatment of GMTV's
dynamic public good game. The pure replication re-analyzes the original
data. \protect\hyperlink{A:-Pure-Replication}{Appendix A} documents
\href{}{errors I identified} in the original paper. The scientific
replication, where I utilize a different sample drawn from a different
population in a different situation, is described in the following
sections.

\hypertarget{sec-design}{%
\subsection{Experimental Design}\label{sec-design}}

As in the NOPUNISH 10 Period treatment of GMTV, I ran sessions with
groups of four (\(i \in I=\{1,2,3,4\}\)), an initial endowment of
\(N_i^1 = 20\) tokens, \(T=10\) periods, a private account with a return
of \(1\) and a group account with a return of \(1.5\) (\(\Rightarrow\)
MPCR\(\equiv \frac{1.5}{4}\)), such that:

\[
N_i^{t+1}=N_i^t - c_i^t + \frac{1.5}{4}\sum_{j=1}^4 c_j^t
\] What makes the game \emph{dynamic}? Instead of receiving fresh
endowments every period, participants received one endowment only at the
beginning of the first period. A participant's endowment in the second
period is the wealth he or she accumulated in the first period. A
participant's endowment in the third period is the wealth he or she
accumulated in the first two periods. And so on. Hence, a decision in
one period has consequences on future endowments and, ultimately, growth
paths. For this reason, the game is described as a \emph{dynamic} public
good game

\hypertarget{sec-sample}{%
\subsection{Sample}\label{sec-sample}}

I recruited the participants from the so called
\emph{\href{https://www.wiso.uni-hamburg.de/forschung/forschungslabor/umfragelabor/aktuelle-umfragen/hamburgpanel.html}{HamburgPanel}}
using HROOT \citep{hroot}. The panel is provided by the University of
Hamburg's Research Laboratory, which used a randomized last digits
approach to build the panel while drawing from the population of
citizens from Hamburg, Germany. Because the sample was exhausted at one
point, I also recruited students from the University of Hamburg.

At the time I conducted the experiment, the representative sample was
not familiar with interactive experiments. In fact, I ran the first
interactive group experiment with this sample. The students, in
contrast, are used to this kind of experiments. Recruiting them, I
excluded those who have participated in a public goods game before. As a
consequence, nonnaiveté is unlikely to affect the validity of my
experiment \citep[ p.~204]{GoodmanPaolacci2017}.

Throughout this paper, I will compare results of my experiment with the
results of GMTV's NOPUNISH 10 Period treatments. I am thus, referring to
three different samples utilized at two points in time: the University
of Nottingham's students (in late 2012), Hamburg's citizens and the
University Hamburg's students (both in July 2021).
Table~\ref{tbl-sample-properties} shows how they compare with respect to
a few properties.

\hypertarget{tbl-sample-properties}{}
\begin{table}[!htbp] \centering 
  \caption{\label{tbl-sample-properties}Sample Properties } 
  \label{} 
\begin{tabular}{@{\extracolsep{5pt}}lcccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{6}{c}{\textit{Dependent variable:}} \\ 
\cline{2-7} 
 & Female & Age & Trust & Meritocracy & Government & Equality \\ 
\hline \\[-1.8ex] 
 Hamburg Citizens & 0.48$^{***}$ & 47.58$^{***}$ & 3.90$^{***}$ & 3.79$^{***}$ & 3.44$^{***}$ & 4.17$^{***}$ \\ 
  & (0.07) & (1.21) & (0.19) & (0.21) & (0.19) & (0.23) \\ 
  & & & & & & \\ 
 Hamburg Students & 0.08 & $-$21.41$^{***}$ & $-$0.58$^{**}$ & 0.23 & 0.71$^{***}$ & 0.12 \\ 
  & (0.09) & (1.63) & (0.25) & (0.28) & (0.26) & (0.30) \\ 
  & & & & & & \\ 
 Nottingham Students & $-$0.10 & $-$15.85$^{***}$ & $-$0.03 & 1.65$^{***}$ & 0.87$^{***}$ & $-$0.35 \\ 
  & (0.09) & (1.52) & (0.23) & (0.26) & (0.24) & (0.28) \\ 
  & & & & & & \\ 
\hline \\[-1.8ex] 
Observations & 208 & 208 & 208 & 208 & 208 & 208 \\ 
R$^{2}$ & 0.02 & 0.47 & 0.04 & 0.20 & 0.06 & 0.02 \\ 
Residual Std. Error & 0.50 & 8.75 & 1.35 & 1.52 & 1.40 & 1.63 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{6}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table}

@@

\hypertarget{sec-software}{%
\subsection{Software}\label{sec-software}}

The experiment was logistically complex for several reasons. First, the
sample was inexperienced. Second, the experiment was interactive and
synchronous. Third, the underlying game was dynamic and interdependent.
This makes dropouts not only more likely, but also more expensive which
is why dropouts were a major concern implementing the experiment.

I chose oTree \citep{oTree} to implement the experiment because it is
open-source, well documented and very flexible. Its
\href{https://getbootstrap.com/}{Bootstrap} (a powerful frontend
toolkit) integration allowed me to make the graphical user interface
interactive, appealing and easy to navigate. The
\href{https://www.highcharts.com/}{Highcharts library} made it easy to
visualize results and to communicate dynamics. The oTree code snippets
made it possible to handle dropouts, for instance, by replacing them
with artificial bots (and informing the rest of the group thereof).
Insofar, oTree served a good tool to consider the participants' user
experience and thus, to make dropouts less likely.

The experiment's source code be found on
\href{https://github.com/Howquez/coopUncertainty}{GitHub}.\footnote{https://github.com/Howquez/coopUncertainty}

\hypertarget{sec-procedure}{%
\subsection{Procedure}\label{sec-procedure}}

Participants entered the experiment at appointed times remotely from
home. They first saw a welcome screen. After agreeing to the privacy
policy, they could proceed to the instructions individually. Having read
these instructions, each participant has also seen a demo-screen
explaining the user interface. Before proceeding, they had to answer six
comprehension questions correctly. Subsequently, they saw a waiting
screen until they could be matched with three other participants, who
have answered the comprehension questions correctly. Once matched, they
were exposed to the decision screen over ten periods. At the end of the
last period, participants saw results of all periods. Subsequently, they
were exposed to a voluntary climate action, where they could donate
(some of) their earnings to offset carbon dioxide. Subsequently, I
elicited risk preferences \citep{HoltLaury2002} and finished with GMTV's
questionnaire.

While I tried to stick to GMTV's protocol as close as possible, I
deviated in a few aspects. First, the instructions were German and also
covered topics inherent to the online setting (dropouts and bots, for
instance). Second, I used another software (oTree instead of zTree).
Third, GMTV gave participants the opportunity to donate to \emph{Doctors
without Borders} whereas we offered carbon dioxide offsets. Fourth, the
graphical user interface looked different.

\hypertarget{sec-results}{%
\section{Results}\label{sec-results}}

\hypertarget{sec-replication}{%
\subsection{Pre-registered GMTV Replication}\label{sec-replication}}

The analyses in this section were pre-registered
\citep{preregistration}. I also pre-registered a RMarkdown file with the
exact code I intended to use
\href{https://github.com/Howquez/coopUncertainty/blob/July21Replication/analysis/reports/rmd}{here}.\footnote{https://github.com/Howquez/coopUncertainty/tree/July21Replication/analysis/reports/rmd
  to run the code, you need to executed the .Rmd files in this
  repository in the order that is indicated by its file names.} Even
though I had to adjust a few lines of code, the main findings in the
following section can also be replicated using the pre-registered code.

\hypertarget{sec-contributions}{%
\subsubsection{Contribution Behavior}\label{sec-contributions}}

First, I ask whether the samples differ with respect to their initial
contributions to the public good. Is our replication sample more
pro-social than the original sample? Figure~\ref{fig-first-round}
reveals that it is not. The distributions of both samples look fairly
similar. Both samples contributed 10 tokens, that is, 50\% of their
endowments on average (median and mean).\footnote{The two-sided rank sum
  test (comparing differences between samples) yields a p-Value of
  0.3926 for the mean contribution in first round of the game.}
Moreover, both samples' initial contributions resemble initial
contributions participants usually make in the standard game with
partner matching.\footnote{See Figure 3B in \citet{fehrgaechter2000}
  (p.989), for instance.} However, in the dynamic game presented here,
we are particularly interested in the subsequent periods because
differences add up exponentially. Do the two groups remain similar over
the course of time?

@@

{\centering \includegraphics{paper_files/figure-pdf/fig-first-round-1.pdf}

@@

\caption{\label{fig-first-round}Individual contributions to the dynamic
public good in the first period}

@@

In particular, do the two samples' contributions follow the same path
over the 10 periods they played? The answer is \emph{no}.
Figure~\ref{fig-share-of-contributions} illustrates that the samples
make similar contributions at the beginning and the end of the game but
behave differently in between. More precisely, the left panel--depicting
the average contributions in absolute terms--shows that the original
sample contributed substantially more than the replication sample
\emph{in all but the first and last period}. For this reason, the
original sample's behavior differs from the replication sample's
behavior in two aspects: it contributes more and exhibits a considerable
drop in the last period (whereas the replication sample's contributions
flatten).

Note that increasing contributions over time imply increasing endowments
over time. Hence, absolute contributions do not us much about the
willingness to cooperate. For this reason, the right panel in
Figure~\ref{fig-share-of-contributions} shows the average \emph{share of
endowments contributed} over time. Both samples exhibit a similar
pattern: their share of endowments contributed declined and did not
stabilize. However, both samples also differ with respect to one aspect:
the replication sample's share of contributions declines faster.

@@

{\centering \includegraphics{paper_files/figure-pdf/fig-share-of-contributions-1.pdf}

@@

\caption{\label{fig-share-of-contributions}The average amount of tokens
contributed over time in treatments.}

@@

\hypertarget{sec-wealth}{%
\subsubsection{Wealth Creation}\label{sec-wealth}}

How do the different contribution-paths translate into
wealth?\footnote{To measure wealth and growth, I define a variable
  called \emph{stock} which sums the endowments of all participants in a
  given group at the end of the round (that is, after the contributions
  have been made, multiplied and redistributed).} Given that the
original sample contributed more in most of the periods, one would
expect the respective groups to be considerably more wealthy.
Figure~\ref{fig-stock-distribution} indicates just that. The grey lines
show that an average group in the original sample accumulated about 478
tokens. In contrast, an average group in the replication sample
accumulated about 380 tokens. This difference is insignificant at
conventional levels\footnote{The two-sided rank sum test (comparing
  differences between samples) yields a p-Value of 0.1356 for the mean
  stock in last round of the game.} though.

@@

{\centering \includegraphics{paper_files/figure-pdf/fig-stock-distribution-1.pdf}

@@

\caption{\label{fig-stock-distribution}Groups' income at the end of the
game}

@@

Although there clearly is growth, groups do not realize the maximal
potential efficiency: under full cooperation, a group can accumulate at
least 4613 tokens or EUR 230. This is depicted in the left panel of
Figure~\ref{fig-growth-heterogeneity}, where one can see the average
wealth over time by sample. The panel illustrates for both samples that
growth was continuous and surprisingly linear, given the exponential
character of the game's design. To sum up, the contribution behavior
differed between samples. In contrast, neither the eventual wealth nor
the corresponding growth paths differed. Differences in contribution
behavior did, thus, not translate to significantly different wealth
outcomes.

Why? Perhaps because the heterogeneity within samples and across groups
has been too large to \emph{detect} a significant difference. The right
panel of Figure~\ref{fig-growth-heterogeneity} depicts heterogeneity: In
the replication sample, the richest group earned 1425 tokens (which is
about 1781\% of the initial endowment) whereas the poorest group ends up
with 92 tokens (115\%). More broadly, the replication sample is
characterized by inequality between groups (\(SD_{Replication} =\)
336.06). The same holds true for the original sample
(\(SD_{Original} =\) 393.58). Hence, the heterogeneity across groups
does not differ between samples, which is remarkable because the
replication sample was drawn from a more heterogeneous (non-convenience
sample). Does it differ within groups?

@@

{\centering \includegraphics{paper_files/figure-pdf/fig-growth-heterogeneity-1.pdf}

@@

\caption{\label{fig-growth-heterogeneity}Average wealth over time across
samples.}

@@

\hypertarget{sec-inequality}{%
\subsubsection{Inequality}\label{sec-inequality}}

Given the different samples and the possibility of endogenous
growth--which essentially is the main feature of the game--I ask whether
and how the inequality grows \emph{within} groups.
Figure~\ref{fig-gini-distribution} illustrates that inequality did grow:
at the end of the game, the original and the replication groups exhibit
an average Gini coefficient of 0.23 and 0.22, respectively.\footnote{The
  two-sided rank sum test (comparing differences between samples) yields
  a p-Value of 0.6059 for the mean Gini coefficient in last round of the
  game.} Because every participant started with the same initial
endowment (in \emph{Period 0}, so to speak), every group started
equally--with a Gini coefficient equaling zero.

Figure~\ref{fig-ginit-time-series} shows that this initial state of
equality ended with the first period already: both samples exhibit a
stark incline in inequality before the second period started. From then
on, the respective Gini coefficients grew slowly but continuously -- for
both samples.

@@

{\centering \includegraphics{paper_files/figure-pdf/fig-gini-distribution-1.pdf}

@@

\caption{\label{fig-gini-distribution}Groups' Gini coefficients (within
groups) at the end of the game}

@@

{\centering \includegraphics{paper_files/figure-pdf/fig-ginit-time-series-1.pdf}

@@

\caption{\label{fig-ginit-time-series}Average Gini coefficient (within
groups) over time across samples}

@@

\textbf{Result 1.} \emph{The \texttt{NOPUNISH\ 10} treatment of GMTV can
be replicated because the replication data resemble the original data
with respect to initial and final contributions, wealth and growth as
well as inequality.}

This is remarkable given the different sample and language, the
different software and user interface as well as the online setting
during the COVID19 pandemic. The result suggests that, by and large, the
sum of these factors did not affect people's preferences towards
cooperation.

\hypertarget{sec-feasibility}{%
\subsection{Online Feasibility}\label{sec-feasibility}}

How did the participants, who have never participated in an online group
experiment before, cope with the situation? Moreover, did participants
understand the unfamiliar setting they found themselves in? While the
answer to the former question requires more thought, the answer to the
latter simply is \emph{yes}: 67 out of 116 answered with \emph{``yes''}
when I asked them. Another 44 answered with \emph{``rather yes''} while
nobody indicated that he or she did not understand the situation at all.
To analyze how participants coped with the situation, I consider three
additional metrics: selection into the experiment, attrition as well as
the time spent on each page.

I first comment on the selection into the experiment: It was difficult
to recruit the sample. The panel counted 1.209 non-students of which we
were able to recruit 130 participants who finished the experiment---even
though we varied the weekdays and timing of the sessions (which were
conducted during a nation-wide lockdown with home office regime). For
this reason, we also recruited students in the last session which
explains the relatively large number of showups in Table~\ref{tbl-meta}.
Although I intended to refrain from the recruitment of students
initially, this particular sub sample enabled me to investigate the
generalizability of my results as I will discuss in
Section~\ref{sec-generalizability}. Alternatively, I also could have
contacted a market research institute to recruit additional participants
within a week. I refrained from doing so to contrast students and the
general population sample, however.
